### 2016-01-16 paper/results3d.tex
regular update

\section{Three-dimensional disks with beta cooling}\label{3ddisk}
We briefly confirm the above results in 3D disks with vertical
structure. This is expected to weaken gravitational instabilities
because the disk mass is spread across the vertical dimension. It is
possible to incorportate this effect in the previous 2D framework, but this
introduces additional parameters. We discuss this in Appendix
\ref{3dcorr}. It is more direct to solve the 3D eigenvalue problem to
avoid such uncertainties. 
%We will find that while softening the
%self-gravity in 2D can produce qualitatively similar results to 3D, a
%quantitative match is

%{\bf note: fast beta-cooling in 3D destablizes vertical modes (enables
%convection)}

%take k>0 wlog
%consider Gamma=1
%height s.t. rho(zmax) = 0.05 rho(0) avoid very low densities 
\subsection{Numerical procedure}
{\bf move to appendix?}
We use a pseudo-spectral method to solve the set of ordinary
differential equations, Eq. \ref{lin_mass}---\ref{lin_gravity}, on 
the domain $ z\in[0,\zmax]$ with a parity condition at the midplane.
We expand $\bm{U}=[\delta P,\delta\rho,\delta\Phi,
  \delta v_x, \delta 
  v_y]$ in even Chebyshev polynomials,
\begin{align}
  \bm{U}(z) = \sum_{j=1}^N \bm{a}_jT_{2(j-1)}(z/\zmax), 
\end{align}
and the vertical velocity in odd Chebyshev polynomials,
\begin{align}
  \delta v_z(z) = \sum_{j=1}^N b_jT_{2j-1}(z/\zmax),   
\end{align}
where $\bm{a}_j$ and $b_j$ are the spectral coefficients. 
The basis functions are chosen to satisfy a reflecting boundary
condition at the midplane, %GI has even symmetry across midplane 
\begin{align}
  \bm{U}^\prime(0) = \bm{0}, \quad \delta v_z(0) = 0.
\end{align}
%At the upper disk boundary we also apply reflection,
We explicitly apply a reflecting upper disk boundary, 
\begin{align}
  \delta v_z(\zmax) = \delta v_x^\prime(\zmax) = \delta
  v_y^\prime(\zmax) = 0,
\end{align}
and the potential satisfies
\begin{align}
  \delta \Phi^\prime (\zmax) + k \delta\Phi(\zmax) = 0, 
\end{align}
as derived by \cite{goldreich65a} and used in similar studies  
\citep{kim12,lin14c}. %reflection is standard in direct simulations

We discretize the equations, including upper disk boundary conditions,
over the $N$ positive abscissae of the extrema of $T_{2N-1}$. This
procedure converts the differential equations into a generalized
eigenvalue problem, for which we use the standard matrix package
LAPACK to solve.

In the following examples we consider a vertically isothermal disk
($\Gamma=1$). Then in the viscous case $\alpha$ is vertically
constant (Eq. \ref{alpha_beta_relation}). We use $N=65$  and apply
numerical disk surface such that $\rho(\zmax)=0.05\rho_0$.

%in practice eliminate density with poisson
\subsection{Inviscid disk}

Fig. \ref{3d_inviscid} show growth rates and the most unstable wavenumber
obtained for a 3D inviscid disk with $\qthree=0.71$
($Q=2Q_\mathrm{crit}$) and $\theta=0$. We also plot 2D
results with the 3D correction as described in Appendix
\ref{3dcorr}. The (empirically) chosen value of
$H_\mathrm{sg}=0.64H$ results in a close match between 2D and 3D
growth rates, but the most unstable wavenumber in 3D is somewhat smaller. 
This reflects self-gravity being weakened in the verical 
direction: a larger horizontal scale is required to achieve the same
strength of self-gravity as the 2D case. Similarly, 
choosing $H_\mathrm{sg}=0.53H$ matches the optimum wavenumbers, but
growth rates are over-estimated in 2D.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.5cm 0.23cm
    0.0cm]{figures/3d_inviscid_rates}\\
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0.23cm
    0.88cm]{figures/3d_inviscid_kmax}
  \caption{Growth rates (top) and optimal wavenumber (bottom) obtained
    from the 3D eigenvalue problem (solid line). Asterisks are
    corresponding values from the 2D dispersion relation
    (Eq. \ref{thindisk}) but with a modified gravitational constant
    as described in Appendix \ref{3dcorr}. \label{3d_inviscid}}
\end{figure}

\subsection{Viscous disk} %growth rates, kmax, eigenfunction
For the 3D viscous problem we use the same set up as that in 2D
(\S\ref{2dvisc}), but with 
\begin{align*}
  \qvert = \frac{Q_\mathrm{3D,crit}}{\sqrt{\alpha}}, 
\end{align*}
where $Q_\mathrm{3D,crit}\simeq0.36$ is the 3D equivalent to the 2D
critical value, $Q_\mathrm{crit}$.

Fig. \ref{3d_visc} show growth rates, maximized over $k$, as a
function of the cooling time $\beta$. We also plot 2D results with 3D
corrections. Softening the self-gravity in 2D captures the correct 
qualitative behavior as in the full 3D case, but it is clear that a
single, constant value of $H_\mathrm{sg}$ cannot re-produce 3D growth 
rates for all $\beta$. 
This suggests that the exact value of
$H_\mathrm{sg}$ is problem-dependent, although taking
$H_\mathrm{sg}\sim O(H)$ should give the correct 3D growth rate within
a factor of two. 
%A smaller $H_\mathrm{sg}$ is needed with faster 
%cooling (larger viscosity) and vice versa. 
%because most unstable wavenumber is smaller with high viscosity

%smaller scales with increasing beta

%, because viscosity makes the
%disk behave more two-dimensional.

%by eliminating vertical velocity gradients.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.cm 0.23cm
    0.0cm]{figures/growth_visc3d}
  \caption{Growth rates from the viscous 3D eigenvalue problem (black solid
    line). The other curves are obtained from the 2D dispersion
    relation (Eq. \ref{thindisk}) but with a modified gravitational
    constant as described in Appendix \ref{3dcorr}. \label{3d_visc}}
\end{figure}