### 2016-01-22 paper/paper3d.tex
re-writing intro

%\documentclass[12pt]{aastex}
\documentclass[iop, numberedappendix]{emulateapj}
%\documentclass{article}
%\usepackage{draftwatermark}
%\usepackage{paralist}
\shorttitle{Non-ideal Gravitational instability}
\shortauthors{M-K.\ Lin \& K. Kratter}

\usepackage{amsmath}
\usepackage{paralist}
%\usepackage{url}
\usepackage{bm}
%\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\newcommand{\p}{\partial}
\newcommand{\zmax}{z_\mathrm{max}}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\sbar}{\bar{\sigma}} 
\newcommand{\He}{\operatorname{He}}
\newcommand{\imgi}{\mathrm{i}} 
\newcommand{\ii}{\mathrm{i}}
\newcommand{\C}{\mathcal{C}^\lambda}
\newcommand{\wop}[1]{\mathcal{V}^{(#1)}}
\newcommand{\avg}[1]{\langle{#1}\rangle}
%\newcommand{\max}[1]{\mathrm{max}({#1})}
%\newcommand{\min}[1]{\mathrm{min}({#1})}
\newcommand{\zhat}{\hat{z}}
\newcommand{\khat}{\hat{k}}
\newcommand{\rhat}{\hat{r}}
\newcommand{\qthree}{Q_\mathrm{3D}}
\newcommand{\qtwo}{Q}
\newcommand{\he}{\mathcal{H}}
\newcommand{\amp}{\mathcal{A}}
\newcommand{\dd}{\delta}
\newcommand{\real}{\operatorname{Re}}
\newcommand{\imag}{\operatorname{Im}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\wbar}{\tilde{W}}
\newcommand{\qbar}{\tilde{Q}}
\newcommand{\dbigH}{\frac{H^\prime}{H}}
\newcommand{\Rp}{\frac{\rho_0^\prime}{\rho_0}}
\newcommand{\Fp}{\frac{g^\prime}{g}}
\newcommand{\Rpp}{\frac{\rho_0^{\prime\prime}}{\rho_0}}
\newcommand{\Fpp}{\frac{g^{\prime\prime}}{g}}
\newcommand{\ils}{L_s^{-1}}
\newcommand{\ilp}{L_p^{-1}}
\newcommand{\ihp}{H_p^{-1}}
\newcommand{\ihs}{H_s^{-1}}
\newcommand{\qcrit}{Q_\mathrm{crit}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\csmid}{c_{s0}} 
\newcommand{\hiso}{h_\mathrm{iso}}
\newcommand{\Hiso}{H_\mathrm{iso}}
\newcommand{\zeus}{\texttt{ZEUS-MP }}
\newcommand{\rsph}{r_\mathrm{sph}}
\newcommand{\dvx}{\dd v_x}
\newcommand{\dvy}{\dd v_y}
\newcommand{\dvz}{\dd v_z}
\newcommand{\dbx}{\dd B_x}
\newcommand{\dby}{\dd B_y}
\newcommand{\dbz}{\dd B_z}
\newcommand{\tcool}{t_\mathrm{c}}
\newcommand{\tirr}{T_\mathrm{irr}}
\newcommand{\w}{ \widetilde{W}}
\newcommand{\dphi}{\dd \Phi}
\newcommand{\hsg}{H_\mathrm{sg}}
\newcommand{\yr}{\mathrm{yr}}
\newcommand{\qvert}{Q_\mathrm{3D}}
\defcitealias{lin12}{L12} 
\defcitealias{nelson13}{N13}
\defcitealias{barker15}{BL15}
\defcitealias{mcnally14}{MP14}
\defcitealias{goldreich67}{GS67}
\defcitealias{rafikov15}{R15}

\def \OmK {\Omega_{\rm K}}

\begin{document}

%\title{untitled linear problem}
\title{On the Gravitational stability of gravito-turbulent accretion disks}
\author{Min-Kai Lin\altaffilmark{1} \& Kaitlin M. Kratter}
\affil{Department of Astronomy and Steward Observatory, \\
University of
  Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA}
\altaffiltext{1}{Steward Theory Fellow}
\email{minkailin@email.arizona.edu}

\begin{abstract}
Low mass, self-gravitating accretion disks admit quasi-steady,
`gravito-turbulent' states in which cooling balances turbulent
heating. However, numerical simulations show that gravito-turbulence
cannot be sustained beyond dynamical timescales when the cooling rate
or turbulent viscosity is too large, resulting in disk fragmentation.        
%We quantify the interpretation of this result as the
%compromise of gravitational stability by: 
We quantify the interpretation of this result as a 
secondary gravitational instability due to: 
\begin{inparaenum}[1)] 
\item 
  cooling, which reduces pressure support; and/or
\item 
  viscosity, which reduces rotational support. 
\end{inparaenum}
We analyze the gravitational stability of accretion disks 
including these non-ideal effects explicitly.    
We derive a general dispersion relation for linear waves in a 
self-gravitating, viscous, non-adiabatic disk with internal and
external heating, and cooling. We 
consider parameterized cooling functions in two and three dimensions,
as well as full radiative diffusion in 3D. 
We apply linear theory to 
global models of protoplanetary disks (PPDs) subject to
gravito-turbulence modeled as an effective 
viscosity and cooling via dust opacity. We find the sole effect of cooling is insufficient to
destabilize a realistic PPD. However, inclusion of viscosity renders
the disk beyond $\sim 60$AU dynamically unstable on radial
%this is for Mdot = 1d-6  
lengthscales a few times the local disk thickness. This is coincident 
with the empirical condition for disk fragmentation based on a
maximum sustainable stress. 
We suggest turbulent stresses play an active role in realistic disk
fragmentation by removing rotational stabilization against 
self-gravity. 
\end{abstract}

\section{Introduction}\label{intro} 
The gravitational instability (GI) of rotating disks    
plays a fundamental role in many astrophysical problems. In the context of 
gaseous protoplanetary disks (PPDs), GI has two applications. It can  
provide gravitational torques to transport angular momentum outwards
and thus enable mass accretion \citep{armitage10,turner14}.   
GI may also lead to disk fragmentation, which has been invoked to 
explain giant planet formation at large radii
\citep{helled14}. Studying these non-linear phenomena require direct      
numerical simulations. Nevertheless, physical insight can be  
obtained through analytical understanding of GI.

The standard metric for the (inverse) strength of disk self-gravity is
the \emph{Toomre parameter}, 
\begin{align}\label{toomreQ_criterion}
  Q \equiv \frac{c_s\kappa}{\pi G \Sigma}  
\end{align}
\citep{toomre64}. Here, $c_s$ is the sound-speed,
$\kappa$ is the epicyclic frequency (which equals the rotation
frequency $\Omega$ in a Keplerian disk), $\Sigma$ is the surface
density and $G$ is the gravitational constant. The Toomre $Q$ is a
measure of the destabilizing effect of self-gravity ($G\Sigma$) against 
the stabilzing effect of rotation ($\kappa$) and pressure 
($c_s$). This is evident from the dispersion relation, 
\begin{align}\label{classic_dr}
  s^2 =  2\pi G\Sigma|k| - \kappa^2 - c_s^2k^2,   
\end{align}
which relates the growth rate $s$ and radial wavenumber $k$ for local,
axisymmetric waves in a two-dimensional (2D, razor-thin), inviscid and
adiabatic (including isothermal) disk. Such disturbances are formally
unstable when the Toomre condition, $Q<1$, is satisfied.
Non-axisymmetric modes can develop for  larger but still order-unity
values of $Q$ . {\bf ref} 
%strictly speaking non axi *local* modes are stable -> only transient
%growth
%true non-axi instability needs global disk  model 
%non-axi GI at slightly larger Q, but still characterized by Q

%qualitative intro to gravito-turbulence 
%role of cooling and resulting viscosity

It is important to remember that the dispersion relation
Eq. \ref{classic_dr}  is derived from \emph{idealized} conditions: 
the disk is laminar and does not experience any net thermal
losses\footnote{Isothermal disks implicitly assume  
  heating and thermal losses are exactly balanced at all 
  times, so these do not cool in the current context.} 
(hereafter `cooling').  
However, real disks can cool, for example,
due to radiative losses. Accretion disks may also be turbulent, for
which its dynamic and thermodynamic effect is often modeled through an
effective viscosity \citep{shakura73,lin87,rafikov15}. {\bf 
  more refs here? non-sg refs} 
%in terms of AM transport and heating 
%assess GI in general

Despite this, the Toomre parameter, and its associated
dispersion relation, are still widely applied to realistic disk models
with viscosity and cooling \citep[e.g.][]{cossins09, kimura12}, 
even though Eq. \ref{toomreQ_criterion}---\ref{classic_dr} were 
derived from inviscid disks without cooling.

However, as we emphasize in this work, the inclusion of non-ideal
physics, such as cooling and viscosity, modifies the dispersion
relation, Eq. \ref{classic_dr}, and hence the condition for GI. This 
means that the Toomre condition alone is insufficient to assess GI
in realistic disks. The goal of this work is to generalize the
analytic treatment of disk GI to allow a self-consistent discussion of
GI in realistic accretion disk models. Specifically, we consider the non-ideal
effects of viscosity and cooling on GI.

%Specifically, we will examine the effect of cooling and  
%viscosity on GI in PPDs.

\subsection{Cooling-driven gravitational instability}
Cooling reduces pressure (thermal) support against self-gravity. The cooling time 
$\tcool$ is the time-scale over which the disk temperature $T$ is relaxed
to some pre-defined value $T_0$, which may be zero. That is, $\p_t T =
-(T-T_0)/\tcool$. It is often written as 
\begin{align} 
  \tcool = \beta \Omega^{-1},
\end{align}
where $\beta$ is the corresponding dimensionless cooling time. This
type of cooling function, first applied by \cite{gammie01}, is a
simple parameterization that allows the full range of thermodynamic
response to be explored. 
{\bf physical origin of cooling in PPDs}

%The Toomre $Q$ parameter does not include cooling, and thus cannot be
%formally applied to assess GI in realistic disks. Indeed, there exists
%a separate, \emph{empirical} cooling time criterion which governs
%stability against disk fragmentation (see below), but this is not well
%understood.

We shall quantify how cooling removes pressure support to enable GI
even when $Q>1$.

We will find that cooling in fact changes the 
qualitative nature of GI, in that its mere presence can
formally lead to instability.

\subsection{Viscosity-driven gravitational instability}
%there is also viscous damping which stabilizes
Viscous disks can develop GI even when $Q>1$. This is because
viscosity compromises rotational support against self-gravity for
long-wavelength disturbances
\citep{lynden-bell74,willerding92,gammie96}. A similar effect  
occurs in dusty fluids where the required frictional forces are 
provided by dust-gas drag \citep{ward00, takahashi14}. In fact, this
is a mechanism to enhance particle clumping for planetesimational formation 
%frag of dust 
\citep{youdin11}. Then it is not unreasonable to expect an analogous
fragmentation of gaseous disks due to viscosity.

For gaseous accretion disks, a Navier-Stokes viscosity is often used to 
to mimic hydrodynamic or magneto-hydrodynamic turbulence
\citep{shakura73}. This involves writing the
kinematic viscosity $\nu$ as  
\begin{align}\label{alpha_param}
  \nu  = \alpha \frac{c_s^2}{\Omega}, %dimensional argument 
\end{align}
where $\alpha$ is the dimensionless viscosity
coefficient. This
parameterization can be modified to include more complex dependencies on the
fluid variables, but Eq. \ref{alpha_param} is the general form.

How well turbulence can be modeled as an effective viscosity is a
separate issue \citep{balbus99}. However, it is  
reasonable to assume that turbulence frustates rotational 
support. (This is, after all, the physical cause for turbulent mass
accretion.) In this work, we use viscosity to represent this physical
effect.

Viscosity is absent in the standard Toomre parameter. The above
studies account for viscosity but consider isothermal disks. On the
other hand, proper inclusion of viscosity should also account for
viscous heating which, as discussed above, also affect stability. 
We thus extend the study of viscous GI to include an energy equation
with viscous heating and explicit cooling. This allows us to consider
GI in realistic PPDs.

%effect of viscosity 
%secular instability 
%but adding viscosity means viscous heating should also be included

\subsection{Relevance to gravito-turbulent disk fragmentation}

% Then the disk temperature will be lowered 
% until $Q\sim 1$, whence non-axisymmetric modes grow 
% and heats the disk through dissipation of spiral shocks. This may 
% lead to a quasi-steady turbulent state with $Q\sim 1$ in
% which cooling is balanced by shock heating to maintain thermal
% equilibrium \citep{gammie01}.

The two effects described above are immediately relevant to
`gravito-turbulence'. This is a steady disk state in which cooling is
balanced by turbulent heating, itself caused by GI. In this case 
the $\alpha$ and $\beta$ parameters are inversely related
\citep{gammie01}.

Experimentally, however, it is found that if $\beta$ ($\alpha$) is
too small (large), 
\begin{align}
  \beta < \beta_\mathrm{c} \quad
  \left(\alpha>\alpha_\mathrm{c}\right), 
\end{align}
then the disk fragments \citep{gammie01,rice05,rice11}. 
There is considerable debate on the exact value of $\beta_\mathrm{c}$
and whether or not a critical cooling time can 
be defined at all \citep{meru11,meru12,paardekooper12,hopkins13}. However, it is
generally accepted that \emph{steady, gravito-turbulent disks do not
  exist for sufficiently rapid cooling}.

The non-existence of steady gravito-turbulence for rapid cooling, or
equivalently large viscosity, is intriguing because, as discussed
above, these effects reduce pressure and rotational support 
against self-gravity, respectively.

In this paper, we consider the stability of PPDs subject to 
gravito-turbulence modeled as an effective viscosity. 
We perform standard stability analyses with explicit inclusion of the
relevant physics. 
We will find that the
regions of a gravito-turbulent PPD that is dynamically unstable,
agrees with the current consensus on where, and the conditions under
which, a PPD is expected to fragment. This motivates an alternative 
interpretation of disk fragmentation as the inability to maintain a
stable, gravito-turbulent state due to the action of cooling and
viscosity.

%We do not attempt to model disk fragmentation from first
%principles. Instead, we simply ask: \emph{when does a 
%  gravito-turbulent disk become dynamically unstable?} 
%We will find the answer to be consistent with the current consensus on
%disk fragmentation.

%viscous gi speculated by gammie 01 but now found in those sims. but
%longer sims of the same system by paardekooper eventually fragment

%important to note that this is an experimental results. we refrain
%from using input from numerical simulations in our modeling 
%because we want to explain simulations 
%we will find the regimes where viscous, cooling PPDs are dynamically
%unstbale coincides with where fragmentation occurs 
% we find instability formally possible for long cooling but takes a
% long time

\subsection{Plan}
%plan of this paper 
%aim to analyse GI in realistic disks with minimal input from numerical
%simulation

The basic equations are listed in \S\ref{basic}, along with
descriptions of the disk 
equilibra and the two cooling models 
considered --- parametrized `beta' cooling and radiative diffusion. 
The linear stability problem is defined in
\S\ref{linear}. We present results with beta cooling in
\S\ref{result_2d} and \S\ref{3ddisk} for 2D and 3D disks,
respectively. %key result: generalized dispersion relation 
In
\S\ref{2dppd} we apply linear theory to PPDs with realistic cooling
models, including radiative diffusion in 3D. We find regions in a
PPD that cannot sustain steady gravito-turbulence to coincide with
that expected for disk fragmentation, based on previous empirical
criteria. We summarize our results in \S\ref{summary}, including a
discussion of how our model may aid the physical understanding of  
fragmentation in realistic PPDs. 
% as the inability to sustain gravito-turbulence
%because the associated stress removes rotational support

\input{model3d}
\input{linear3d}
\input{results2d}
\input{results3d}
\input{application} 
\input{summary}

%\acknowlegement
%S. Stahler

\appendix
\input{appendix3d}

\bibliographystyle{apj}
\bibliography{ref}

\end{document}