### 2016-01-22 paper/paper3d.tex
intro version 2

%\documentclass[12pt]{aastex}
\documentclass[iop, numberedappendix]{emulateapj}
%\documentclass{article}
%\usepackage{draftwatermark}
%\usepackage{paralist}
\shorttitle{Non-ideal Gravitational instability}
\shortauthors{M.-K.\ Lin \& K. Kratter}

\usepackage{amsmath}
\usepackage{paralist}
%\usepackage{url}
\usepackage{bm}
%\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\newcommand{\p}{\partial}
\newcommand{\zmax}{z_\mathrm{max}}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\sbar}{\bar{\sigma}} 
\newcommand{\He}{\operatorname{He}}
\newcommand{\imgi}{\mathrm{i}} 
\newcommand{\ii}{\mathrm{i}}
\newcommand{\C}{\mathcal{C}^\lambda}
\newcommand{\wop}[1]{\mathcal{V}^{(#1)}}
\newcommand{\avg}[1]{\langle{#1}\rangle}
%\newcommand{\max}[1]{\mathrm{max}({#1})}
%\newcommand{\min}[1]{\mathrm{min}({#1})}
\newcommand{\zhat}{\hat{z}}
\newcommand{\khat}{\hat{k}}
\newcommand{\rhat}{\hat{r}}
\newcommand{\qthree}{Q_\mathrm{3D}}
\newcommand{\qtwo}{Q}
\newcommand{\he}{\mathcal{H}}
\newcommand{\amp}{\mathcal{A}}
\newcommand{\dd}{\delta}
\newcommand{\real}{\operatorname{Re}}
\newcommand{\imag}{\operatorname{Im}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\wbar}{\tilde{W}}
\newcommand{\qbar}{\tilde{Q}}
\newcommand{\dbigH}{\frac{H^\prime}{H}}
\newcommand{\Rp}{\frac{\rho_0^\prime}{\rho_0}}
\newcommand{\Fp}{\frac{g^\prime}{g}}
\newcommand{\Rpp}{\frac{\rho_0^{\prime\prime}}{\rho_0}}
\newcommand{\Fpp}{\frac{g^{\prime\prime}}{g}}
\newcommand{\ils}{L_s^{-1}}
\newcommand{\ilp}{L_p^{-1}}
\newcommand{\ihp}{H_p^{-1}}
\newcommand{\ihs}{H_s^{-1}}
\newcommand{\qcrit}{Q_\mathrm{crit}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\csmid}{c_{s0}} 
\newcommand{\hiso}{h_\mathrm{iso}}
\newcommand{\Hiso}{H_\mathrm{iso}}
\newcommand{\zeus}{\texttt{ZEUS-MP }}
\newcommand{\rsph}{r_\mathrm{sph}}
\newcommand{\dvx}{\dd v_x}
\newcommand{\dvy}{\dd v_y}
\newcommand{\dvz}{\dd v_z}
\newcommand{\dbx}{\dd B_x}
\newcommand{\dby}{\dd B_y}
\newcommand{\dbz}{\dd B_z}
\newcommand{\tcool}{t_\mathrm{c}}
\newcommand{\tirr}{T_\mathrm{irr}}
\newcommand{\w}{ \widetilde{W}}
\newcommand{\dphi}{\dd \Phi}
\newcommand{\hsg}{H_\mathrm{sg}}
\newcommand{\yr}{\mathrm{yr}}
\newcommand{\qvert}{Q_\mathrm{3D}}
\defcitealias{lin12}{L12} 
\defcitealias{nelson13}{N13}
\defcitealias{barker15}{BL15}
\defcitealias{mcnally14}{MP14}
\defcitealias{goldreich67}{GS67}
\defcitealias{rafikov15}{R15}

\def \OmK {\Omega_{\rm K}}

\begin{document}

%\title{untitled linear problem}
\title{On the Gravitational stability of gravito-turbulent accretion disks}
\author{Min-Kai Lin\altaffilmark{1} \& Kaitlin M. Kratter}
\affil{Department of Astronomy and Steward Observatory, \\
University of
  Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA}
\altaffiltext{1}{Steward Theory Fellow}
\email{minkailin@email.arizona.edu}

\begin{abstract}
Low mass, self-gravitating accretion disks admit quasi-steady,
`gravito-turbulent' states in which cooling balances turbulent
heating. However, numerical simulations show that gravito-turbulence
cannot be sustained beyond dynamical timescales when the cooling rate
or turbulent viscosity is too large, resulting in disk fragmentation.        
% `alternative' or `physical' here?
%quantify or develop
We motivate and quantify an alternative interpretation of disk
fragmentation as the inability to maintain gravito-turbulence due to  
secondary gravitational instabilities driven by:  
\begin{inparaenum}[1)] 
\item 
  cooling, which reduces pressure support; and/or
\item 
  viscosity, which reduces rotational support. 
\end{inparaenum}
We analyze the gravitational stability of accretion disks 
including these non-ideal effects explicitly.    
We derive a general dispersion relation for linear waves in a 
self-gravitating, viscous, non-adiabatic disk with internal heating,
external irradiation, and cooling. We  
consider parameterized cooling functions in two and three dimensions,
as well as full radiative diffusion in 3D. We show that 
cooling/viscosity can destabilize the disk even when the classic
Toomre stability condition is satisfied; and generally there is
no critical cooling rate/viscosity below which the disk is formally
stable. %there is no special/precise cooling/visc that marks the
        %stable/unstable boundary
We apply linear theory to 
global models of protoplanetary disks (PPDs) subject to
gravito-turbulence modeled as an effective 
viscosity and cooling via dust opacity. 
%We find the sole effect of
%cooling is insufficient to 
%destabilize a realistic PPD. 
We find that viscosity renders the PPD beyond $\sim 60$AU dynamically
unstable on radial %this is for Mdot = 1d-6  
lengthscales a few times the local disk thickness. This is coincident 
with the empirical condition for disk fragmentation based on a
maximum sustainable stress. 
We suggest turbulent stresses can play an active role in realistic disk 
fragmentation by removing rotational stabilization against 
self-gravity. {\bf shorten abstract more?}
\end{abstract}

\section{Introduction}\label{intro} 
The gravitational instability (GI) of rotating disks    
plays a fundamental role in many astrophysical problems. In the context of 
gaseous protoplanetary disks (PPDs), GI has two applications. It can  
provide gravitational torques to transport angular momentum outwards
and thus enable mass accretion \citep{armitage10,turner14}.   
GI may also lead to disk fragmentation, which has been invoked to 
explain giant planet formation at large radii
\citep{helled14}. Studying these non-linear phenomena require direct      
numerical simulations. Nevertheless, physical insight can be  
obtained through analytical understanding of GI.

The standard metric for the (inverse) strength of disk self-gravity is
the \emph{Toomre parameter}, 
\begin{align}\label{toomreQ_criterion}
  Q \equiv \frac{c_s\kappa}{\pi G \Sigma}  
\end{align}
\citep{toomre64}. Here, $c_s$ is the sound-speed,
$\kappa$ is the epicyclic frequency (which equals the rotation
frequency $\Omega$ in a Keplerian disk), $\Sigma$ is the surface
density and $G$ is the gravitational constant. The Toomre $Q$ is a
measure of the destabilizing effect of self-gravity ($G\Sigma$) against 
the stabilzing effect of rotation ($\kappa$) and pressure 
($c_s$). This is evident from the dispersion relation, 
\begin{align}\label{classic_dr}
  s^2 =  2\pi G\Sigma|k| - \kappa^2 - c_s^2k^2,   
\end{align}
which relates the growth rate $s$ and radial wavenumber $k$ for local,
axisymmetric waves in a two-dimensional (2D, razor-thin), inviscid and
adiabatic (including isothermal) disk. Such disturbances are formally
unstable when $Q<1$. Non-axisymmetric modes can develop for larger,
but still order-unity values of $Q$
\citep{papaloizou89,papaloizou91}. 
%strictly speaking non axi *local* modes are stable -> only transient
%growth
%true non-axi instability needs global disk  model 
%non-axi GI at slightly larger Q, but still characterized by Q

It is important to remember that the popular dispersion relation, 
Eq. \ref{classic_dr},  is derived from \emph{idealized} conditions: 
the disk is laminar and does not experience any net thermal
losses\footnote{Isothermal disks implicitly assume  
  heating and thermal losses are exactly balanced at all 
  times, so these do not cool in the current context.} 
(hereafter `cooling').  
However, real disks can cool, for example, %radiative, convective cooling
due to radiative losses. Accretion disks may also be turbulent, for
which its dynamic and thermodynamic effect is often modeled through an
effective viscosity \citep{shakura73,lin87,rafikov15}. {\bf 
  more refs here?} 
%in terms of AM transport and heating 
%assess GI in general

Despite this, the Toomre parameter and its associated
dispersion relation are still widely applied to viscous, cooling
accretion disk models \citep[e.g.][]{cossins09, kimura12}, 
even though Eq. \ref{toomreQ_criterion}---\ref{classic_dr} were 
derived from inviscid disks without cooling.

However, as we emphasize in this work, the inclusion of non-ideal
physics, such as cooling and viscosity, modifies the dispersion
relation, Eq. \ref{classic_dr}, and hence the condition for GI. This 
means that the Toomre condition alone is insufficient to assess GI
in realistic disks. The goal of this work is to generalize the
analytic treatment of disk GI to allow a self-consistent discussion of
GI in realistic accretion disk models. Specifically, we will examine
how  cooling and viscosity can lead to GI even when $Q>1$, by 
removing pressure and rotational support, respectively.

%Specifically, we will examine the effect of cooling and  
%viscosity on GI in PPDs.

\subsection{Cooling-driven gravitational instability}\label{cool_gi}
Cooling reduces pressure support against self-gravity. The
cooling time  
$\tcool$ is the time-scale over which the disk temperature $T$ is relaxed
to some floor value $T_0$, which may be zero. That is, $\p_t T =
-(T-T_0)/\tcool$. It is often written as 
\begin{align} 
  \tcool = \beta \Omega^{-1},
\end{align}
where $\beta$ is the corresponding dimensionless cooling time. This
type of cooling function, first applied by \cite{gammie01}, is a
simple parameterization that allows the full range of thermodynamic
response to be explored. 
{\bf physical origin of cooling in PPDs}

We will quantify how cooling removes pressure support to enable GI
even when $Q>1$. For example, if perturbations are
cooled towards zero tempertaure, we find the above dispersion relation is
modified to read 
\begin{align}\label{cool_dr}
s^2 = 2\pi G\Sigma |k| - \kappa^2 - \gamma \left(\frac{\tcool s}{1 +
  \tcool s}\right)c_s^2k^2,
\end{align}
where $\gamma$ is the adiabatic index. This equation shows that
cooling increases the growth rate by reducing the magnitude of the
pressure term. In fact, Eq. \ref{cool_dr} implies that instability is
possible for any finite $Q$. %any finite pair of Q and beta 
This shows that cooling can change the
qualitative nature of GI compared to the simple Toomre condition.

%current understanding of frag is all due to cooling (viscosity is
%just a reflection of cooling)

\subsection{Viscosity-driven gravitational instability}\label{visc_gi}
%there is also viscous damping which stabilizes
Viscous disks can also develop GI even when $Q>1$. This is because, as 
demonstrated below, viscosity compromises rotational support against  
self-gravity for long-wavelength disturbances 
\citep{lynden-bell74,willerding92,gammie96}. A similar effect   
occurs in dusty fluids where the required frictional forces are 
provided by dust-gas drag \citep{ward00, takahashi14}. In fact, this
is a mechanism to enhance particle clumping for planetesimational
formation %frag of dust 
\citep{youdin11}. Then it is not unreasonable to expect an analogous
fragmentation of gaseous disks due to viscosity.

It is conventional to write the kinematic viscosity $\nu$ as   
\begin{align}\label{alpha_param}
  \nu  = \alpha \frac{c_s^2}{\Omega}, %dimensional argument 
\end{align}
where $\alpha$ is the dimensionless viscosity
coefficient. This
parameterization can be modified to include more complex dependencies on the
fluid variables, but Eq. \ref{alpha_param} is the general form. 
{\bf maybe no need to intro alpha param here?}

For an isothermal, viscous, self-gravitating disk in 2D,
\cite{gammie96} finds the approximate disperion relation   
\begin{align}
s \simeq \frac{\nu k^2\left(2\pi G\Sigma |k| -
  c_s^2k^2\right)}{\kappa^2 + c_s^2k^2 - 2\pi G\Sigma|k|}. 
\end{align}
Assuming $Q>1$, instability occurs if 
\begin{align*}
  2\pi G \Sigma > c_s^2|k|,
\end{align*}
provided that $\nu k^2\neq0$. The above condition for viscous GI is 
identical that would be obtained from Eq. \ref{classic_dr} with
rotation neglected ($\Omega\to 0$). That is, a classically stable
disk can be made unstable by viscosity as it effectively removes
rotational stabilization \citep{lynden-bell74}.

%turbulence <-> viscosity assumption 
For gaseous accretion disks, a Navier-Stokes viscosity is often
implemented (as above) to mimic hydrodynamic or magneto-hydrodynamic
(MHD) turbulence \citep{shakura73}. How well turbulence can be modeled
as an effective viscosity is a separate issue 
\citep{balbus99}. However, it is reasonable to assume that turbulence
frustates rotational support. (This is, after all, the physical
mechanism to turbulent mass accretion.) 
For example, \cite{fromang05} found numerical evidence that MHD
turbulence can enhance GI and lead to disk fragmentation (see also
\S\ref{MHD}).

In this work, we use viscosity to model such a 
physical effect due to turbulence, but we note that viscous 
GI does not assume any origin for the viscosity. We will extend theory
of viscous GI to include an energy equation  with viscous heating,
irradiation, explicit cooling, as well as three-dimensionality. This
allows us to consider viscous GI in more realistic PPD models.

%effect of viscosity 
%secular instability 
%but adding viscosity means viscous heating should also be included

\subsection{Relevance to gravito-turbulent disk fragmentation}
We develop a general framework for viscous disks without  
assuming a specific origin of turbulence that this viscosity 
represents. We will, however, apply the theory to `gravito-turbulent'
disks in which the viscosity associated with some underlying (classic)
GI, as described below.

Consider a disk initially laminar without a heating mechanism and
cools. Then the disk temperature will be lowered until $Q=O(1)$, whence
non-axisymmetric modes grow and heats the disk through the dissipation
of spiral shocks. %new rafikov 16 
This permits a quasi-steady, turbulent state with
$Q =O(1)$ in which cooling is balanced by shock heating to maintain
thermal equilibrium \citep{gammie01}. 
The dynamic and thermodynamic effect of this turbulence will be
represented by a viscosity. Then the $\alpha$ and $\beta$ parameters
defined above will be inversely related.

Experimentally, however, it is found that if the cooling rate or
viscosity is too large, say,   
\begin{align}\label{frag_cond}
  \beta < \beta_\mathrm{c} \quad
  \left(\alpha>\alpha_\mathrm{c}\right), 
\end{align}
then the disk fragments \citep{gammie01,rice05,rice11}. In fact, in
the absence of global effects, numerical simulations show there are
two --- and only two --- possible outcomes for self-gravitating,
cooling disks: gravito-turbulence or fragmentation.  
%more accurately: given a total simulation time, there are two
%outcomes 
%sufficiently long sims -> always frag?

There is considerable debate on the exact value of $\beta_\mathrm{c}$ 
and whether or not a critical cooling time can be defined at all 
\citep{meru11,lodato11,meru12,paardekooper12,hopkins13}. 
%mention numerical effects? 
However, it is generally accepted that \emph{steady, gravito-turbulent
  disks do not  
  exist for sufficiently rapid cooling or large viscosity}.  
This intriguing because, as highlighted in
\S\ref{cool_gi}---\ref{visc_gi}, these effects can reduce
gravitational stability. These considerations motivate a physical    
interpretation of disk fragmentation as the inability to maintain a
gravito-turbulent state due to the action of cooling and/or viscosity.

Indeed, we will find that the region where a gravito-turbulent PPD is
dynamically unstable agrees with the current consensus on  where, and
the conditions under which, a PPD is expected to fragment based on
Eq. \ref{frag_cond}, with typical critical values cited in the
literature \citep[e.g. $\beta_c\sim
  3,\,\alpha_c\sim0.1$;][]{gammie01,rice05}.

\subsection{Plan}
%plan of this paper 
%aim to analyse GI in realistic disks with minimal input from numerical
%simulation

The basic equations, disk equilibria, and the two cooling models  
considered --- parametrized `beta' cooling and radiative diffusion, 
are given in \S\ref{basic}. 
The linear stability problem is defined in
\S\ref{linear}. We present results with beta cooling in
\S\ref{result_2d} and \S\ref{3ddisk} for 2D and 3D disks,
respectively. %key result: generalized dispersion relation 
In \S\ref{2dppd} we consider PPDs with realistic viscosity/cooling   
models, including radiative diffusion in 3D, and determine where such
disks are gravitationally unstable. We summarize our results in
\S\ref{summary} with a discussion of how our models may aid the 
physical understanding of fragmentation in realistic PPDs.

\input{model3d}
\input{linear3d}
\input{results2d}
\input{results3d}
\input{application} 
\input{summary}

%\acknowlegement
%S. Stahler 
%A. Youdin

\appendix
\input{appendix3d}

\bibliographystyle{apj}
\bibliography{ref}

\end{document}