### 2016-01-23 paper/model3d.tex
model version 2

\section{Basic equations}\label{basic}
We consider a 3D, self-gravitating,
viscous disk with heating and cooling. We use the shearing box framework to
study a small patch of the disk \citep{goldreich65}. The local frame
co-rotates with a fiducial point in the unperturbed disk at angular
frequency $\Omega$. The Cartesian co-ordinates $(x,y,z)$ corresponds
to the radial, azimuthal and verical directions in the global disk. 
The fluid equations are
\begin{align}
  &\frac{\p\rho}{\p t} + \nabla\cdot\left(\rho\bm{v}\right) = 0, \\
  & \frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} =
  -\frac{1}{\rho}\nabla P - \nabla\Phi - 2\Omega\hat{\bm{z}}\times\bm{v}\notag\\ & 
  \phantom{\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} =}
  +2\Omega^2qx\hat{\bm{x}} -\Omega_z^2z\hat{\bm{z}} 
   + \frac{1}{\rho}\nabla\cdot\bm{T},\label{momentum_eq}\\
  & \frac{\p E}{\p t} + \nabla\cdot\left(E\bm{v}\right) = - P
  \nabla\cdot\bm{v} + \mathcal{H}_\mathrm{visc} - \Lambda +
  \mathcal{H}_\mathrm{ext} ,\label{energy_eq}
\end{align} 
where $\rho$ is the density field and $\bm{v} = (v_x, v_y, v_z)$ is
the velocity field.% $q$ is the shear parameter and $\Omega_z$ is the
%local vertical oscillation frequency. 
We assume an ideal gas so that the pressure $P$ and thermal energy
density $E$ are related by  
\begin{align}
  P = (\gamma-1)E = \mathcal{R}\rho T, 
\end{align}
where $\mathcal{R}$ is the gas constant and $T$ is the temperature. We
use $\gamma=1.4$ unless otherwise stated. The gas 
gravitatational potential $\Phi$ is given via the Poisson equation,  
\begin{align}
 \nabla^2\Phi = 4 \pi G \rho. 
\end{align}
%where $G$ is the gravitational constant.

In the momentum equation (Eq. \ref{momentum_eq}), the third,
fourth/fifth, and last term on the right-hand side represent the Coriolis, 
tidal, and viscous forces (see below), respectively. We consider    
Keplerian disks with shear paramter $q=3/2$ and vertical oscillation 
frequency $\Omega_z=\Omega$.

In the energy equation (Eq. \ref{energy_eq}) the source terms 
$\mathcal{H}_\mathrm{visc}$ and $\Lambda$
represent viscous heating and time-dependent cooling, 
respectively, and $\mathcal{H}_\mathrm{ext}$ represents any
time-independent heat source/sinks. We normally set
$\mathcal{H}_\mathrm{ext}=~0$ unless otherwise stated.

\subsection{Viscosity and heating}\label{visc_model}
The Cartesian components of the viscous stress tensor $\bm{T}$ is
defined by 
\begin{align}
  T_{ij} \equiv \rho \left[\nu \left(\p_j v_i + \p_i v_j\right) +
    \left(\nu_b-\frac{2}{3}\nu\right)\delta_{ij}\nabla\cdot\bm{v}\right], 
\end{align}
%bulk visc ignored 
where $\rho\nu$ is the shear viscosity. We also include a bulk
viscosity $\rho\nu_b$ for completeness, but will neglect it in
numerical calculations. The associated viscous heating given by 
\begin{align}
  \mathcal{H}_\mathrm{visc} \equiv \left(\p_jv_i\right)T_{ij}, 
\end{align}
where summation over repeated indices is implied.

We adopt the viscosity law  
\begin{align}\label{visc_law}
  \nu = \alpha
  \frac{c_{s0}^2}{\Omega}\left(\frac{\rho}{\rho_\mathrm{eq}}\right)^\mu\left(\frac{P}{P_\mathrm{eq}}\right)^\lambda,             
\end{align}
where subscript `eq' denotes the equilibrium state and  
$c_{s0}^2\equiv P_\mathrm{eq}(z=0)/\rho_\mathrm{eq}(z=0)$. 
The dimensionless viscosity coefficient 
$\alpha=~\alpha(\rho_\mathrm{eq},P_\mathrm{eq})$ characterizes the
magnitude of the shear viscosity \emph{in steady state}. The indices
$\mu,\,\lambda$ are free parameters chosen to model how the viscosity
behaves in the perturbed state.  
We adopt the same prescription for the bulk viscosity but
with $\alpha\to\alpha_b$.

Our numerical calculations use $\mu=-1,\lambda=0$ so 
that $\rho\nu$ is time-indpendent, folling previous 
studies of viscous GI  
\citep{lynden-bell74,hunter83,willerding92,gammie96}.  
This choice also conveniently eliminates viscous over-stability
\citep{schmit95,latter06}, which is unrelated to self-gravity, that
would otherwise contaminate our results. %obscure

While steady-state viscosity values can be determined 
analytically \citep[e.g.][]{rafikov15}, its time-dependent behavior is
not well-explored. We emphasize the choice $\mu=-1,\lambda=0$ is made
to bring out the physical process of interest --- viscous GI. %expect
                                %to 
%be generally present, but will 
%be obscured by viscous over-stability with
%other indices
Interestingly, though, \cite{laughlin96b} have suggested a $\nu \propto 1/\Sigma$
dependence when modeling the evoluting of 2D self-gravitating disks 
as a viscous process. We note that if $\rho\nu$ is constant in time
then increasing the density leads to reduction in the viscosity. 
This is perhaps not inconsistent with numerical simulations of disk
fragmentation which show the internal structure of clumps, which
represent increased density, is laminar \citep{gammie01}.  
{\bf is this true?}

%Furthermore, \cite{paardekooper12} measured
%an increase in the effective viscosity when their gravito-turbulent
%disk models underwent fragmentation (i.e. as $\Sigma$ increases).  
%{\bf note paardekooper 2012: alpha goes down as Sigma goes up}

\subsection{Steady states and cooling models}
We consider equilibrium solutions (here omitting the `eq' subscripts
for simplicity)  
\begin{align}
  \bm{v} &= -q\Omega x \hat{\bm{y}}, \\
  \rho   &= \rho(z),\\
  P      &= P(z) \equiv c_s^2(z)\rho.
\end{align} 
The equilibrium density and pressure fields are obtained by solving
the vertical momentum equation with self-gravity,
\begin{align}
  &\frac{1}{\rho}\frac{dP}{dz} +
  \Omega_z^2z + \frac{d\Phi}{dz} = 0, \label{vert_eq1}\\
 &\frac{d^2\Phi}{dz^2} = 4 \pi G \rho,\label{vert_eq2}
\end{align}
together with thermodynamic equilibrium,
\begin{align}\label{thermal_eq}
(q\Omega)^2\rho\nu + \mathcal{H}_\mathrm{ext} = \Lambda,
\end{align}
where the first term represents viscous heating. For the viscous
problem, $\nu\neq0$, we may set $\mathcal{H}_\mathrm{ext}=0$. However,
if we wish to neglect viscosity but include cooling, we must invoke
$\mathcal{H}_\mathrm{ext}\neq0$ to define an equilibrium state.  
To proceed further, we separately describe the two cooling models
considered in this work. 
%It is possible to have both models present - but too complicated

\subsubsection{Beta cooling}
In this case the energy loss per unit volume is specified 
as an explicit function of the thermodynamic variables. 
A prototypical example is 
\begin{align}\label{beta_cool}
  \Lambda =
  %  \frac{P}{(\gamma-1)T}\frac{\left(T-T_\mathrm{irr}\right)}{t_c}, 
  \frac{\mathcal{R}\rho}{(\gamma-1)}\frac{\left(T-T_\mathrm{irr}\right)}{t_c}. 
\end{align}
Recall from Eq. \ref{beta_def} that 
%\begin{align}
  $\tcool = \beta\Omega^{-1}$ is 
%\end{align}
the cooling timescale with $\beta$ a constant input parameter. Here, 
\begin{align}
  \tirr = \theta T_\mathrm{eq}
\end{align}
is a reference temperature field parameterized as a constant 
fraction $\theta$ of the equilibrium temperature
$T_\mathrm{eq}$. Physically, $\tirr$ is the floor temperature set by,
for example, stellar irradiation. We will thus refer to $\theta$ as
the `irradiation level'. %stellar illum or env rad

We will formulate the linear problem with beta cooling using the
functional form for $\Lambda$ as given by
Eq. \ref{beta_cool}. However, any explicit cooling function, say
$\Lambda=\Lambda_1(\rho,T)$, can be mapped to that obtained from
Eq. \ref{beta_cool} by equating their linearized forms, i.e. $\delta
\Lambda \equiv \delta \Lambda_1$, which then defines the $\beta$ and
$\theta$ parameters to be used in the framework we develop. This is
done in \S\ref{2dppd} where we adopt a more realistic beta cooling 
prescription for 2D PPDs. In this case, $\theta$ may or may not
represent a physical irradiation.

For models with beta cooling given by Eq. \ref{beta_cool} we assume an 
equilibrium polytropic relation 
\begin{align} 
  P  =
%  \frac{c_{s0}^2\rho_0}{\Gamma}\left(\frac{\rho}{\rho_0}\right)^\Gamma,
  %  \equiv \frac{c_s^2(z)}{\Gamma}\rho,
c_{s0}^2\rho_0\left(\frac{\rho}{\rho_0}\right)^\Gamma,
\end{align}
where $\Gamma$ is the constant polytropic index and $\rho_0 =
\rho(z=0)$ is the equilibrium midplane density. 
The vertical structure is first obtained from
Eq. \ref{vert_eq1}---\ref{vert_eq2} then inserted into
Eq. \ref{thermal_eq} to infer the required viscosity profile for
thermal equilibrium,   
\begin{align}\label{alpha_beta_relation}
\alpha(z) = \frac{1}{(\gamma-1)\beta
   q^2}\frac{c_s^2(z)}{c_{s0}^2}\left(1 - \theta\right). 
%  \alpha(z) = \frac{1}{\Gamma(\gamma-1)\beta
%    q^2}\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left(1 - \theta\right). 
\end{align}

Beta cooling is widely applied in 2D and 3D
numerical simulations of self-gravitating disks \citep{gammie01,
  rice05,baehr15}. It permits numerical experiments to be carried out in a
controlled manner as a function of the cooling time $\beta$. An
adiabatic disk corresponds to $\beta\to \infty$. The
physical meaning of the limit $\beta\to0$ depends on $\theta$, as
discussed in \S\ref{theta0}---\ref{theta1}.

\subsubsection{Radiative cooling}\label{rad_cool}
Here we assume energy transfer by radiative diffusion. Then
\begin{align}
  \Lambda &= \nabla\cdot\bm{F},\label{rad_cool1}\\
  \bm{F}   &= -\frac{16\sigma T^3}{3\kappa_d\rho}\nabla T, \label{rad_cool2}
\end{align}
where $\sigma$ is the Stefan-Boltzmann constant and 
$\kappa_d$ is the (dust) opacity. We adopt
\begin{align}\label{opacity_law}
  \kappa_d = \kappa_{d0}T^b,
\end{align}
and take the constant index $b=2$ as appropriate for the cold outer regions
of a realistic PPD \citep{bell94}, but retain the general notation $b$
to keep track of the opacity.    
%We use radiative cooling when considering 3D PPDs in \S.

In this case, we specify a constant viscosity coefficient $\alpha$ and
solve Eq. \ref{vert_eq1}---\ref{thermal_eq}, together with
Eq. \ref{rad_cool1}---\ref{rad_cool2}, as a fourth order system of
ordinary differential equations to obtain equilibrium profiles $P(z)$, $T(z)$, and hence
$\rho(z)$.

While radiative cooling is arguably more realistic than
parametrized beta cooling, it generally implies a vertically
non-isothermal equilibrium disk, and increases the order of the
linearized equations. It formally applies to optically-thick 
disks, but it is possible to modify the flux function to account
for optically-thin disks. However, this complication is beyond the scope of
this work.