### 2016-01-24 paper/results3d.tex
ready to revise applications

\section{Three-dimensional disks with beta cooling}\label{3ddisk}
We confirm the above results in 3D disks with vertical
structure. This is expected to weaken gravitational instabilities
because the disk mass is spread across the vertical dimension. It is
possible to incorportate this effect in the previous 2D framework, but this
introduces additional parameters as discussed in Appendix
\ref{3dcorr}. It is more direct to solve the 3D eigenvalue problem to
avoid such uncertainties. Our numerical approach is outlined in Appendix
\ref{3d_method}.

%We will find that while softening the
%self-gravity in 2D can produce qualitatively similar results to 3D, a
%quantitative match is  
%{\bf note: fast beta-cooling in 3D destablizes vertical modes (enables
%convection)}
%take k>0 wlog
%consider Gamma=1
%height s.t. rho(zmax) = 0.05 rho(0) avoid very low densities

In the following examples we consider a vertically isothermal disk
($\Gamma=1$). Then in the viscous case $\alpha$ is vertically
constant (Eq. \ref{alpha_beta_relation}). We consider even modes about
the midplane, ane apply a numerical disk surface at $z=\zmax$ such
that $\rho(\zmax)=0.05\rho_0$.

%in practice eliminate density with poisson
\subsection{Inviscid disk}

We consider an inviscid disk with $\qthree=0.71$ and $\theta=0$. 
Recall the 3D gravity parameter $\qthree$ is defined by 
Eq. \ref{Q3d_def}. This corresponds to a Toomre parameter of
$Q=2Q_\mathrm{crit}$; where $Q_\mathrm{crit}$, defined by
Eq. \ref{qcrit_def}, is the Toomre parameter value such that the 2D
disk would be marginally stable in the absence of cooling.

Fig. \ref{3d_inviscid} show growth rates and the most unstable
wavenumber obtained for this model. We also plot 2D
results with the 3D correction as described in Appendix
\ref{3dcorr}. The (empirically) chosen value of
$H_\mathrm{sg}=0.64H$ results in a close match between 2D and 3D
growth rates, but the most unstable wavenumber in 3D is somewhat smaller. 
This reflects self-gravity being weakened in the vertical 
direction: a larger horizontal scale is required to achieve the same
strength of self-gravity as the 2D case. Similarly, 
choosing $H_\mathrm{sg}=0.53H$ matches the optimum wavenumbers, but
growth rates are over-estimated in 2D.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.5cm 0.23cm
    0.0cm]{figures/3d_inviscid_rates}\\
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0.23cm
    0.88cm]{figures/3d_inviscid_kmax}
  \caption{Growth rates (top) and optimal wavenumber (bottom) obtained
    from the 3D eigenvalue problem (solid line). Asterisks are
    corresponding values from the 2D dispersion relation
    (Eq. \ref{thindisk}) but with a modified gravitational constant
    as described in Appendix \ref{3dcorr}. \label{3d_inviscid}}
\end{figure}

\subsection{Viscous disk} %growth rates, kmax, eigenfunction
For the 3D viscous problem we use the same set up as that in 2D
(\S\ref{2dvisc}), but with 
\begin{align*}
  \qvert = \frac{Q_\mathrm{3D,crit}}{\sqrt{\alpha}}, 
\end{align*}
where $Q_\mathrm{3D,crit}\simeq0.36$ is the 3D equivalent to the 2D
critical value, $Q_\mathrm{crit}$.

Fig. \ref{3d_visc} show growth rates, maximized over $k$, as a
function of the cooling time $\beta$. We also plot 2D results with 3D
corrections. Softening the self-gravity in 2D captures the correct 
qualitative behavior as in the full 3D case, but it is clear that a
single, constant value of $H_\mathrm{sg}$ cannot re-produce 3D growth 
rates for all $\beta$. 
This suggests that the exact value of
$H_\mathrm{sg}$ is problem-dependent, although taking
$H_\mathrm{sg}\sim O(H)$ should give the correct 3D growth rate within
a factor of two.

%cooling helps vertical modes (convection) 
%slower variation wrt z for low beta because of associated high
%viscosity
%smooth out vertical gradients 
Fig. \ref{3d_visc_vz} shows the magnitude of the vertical velocity
$|\delta v_z|$ scaled by the total horizontal velocity for $\beta =
1,\,10$ and $100$. Cooling generally enhances vertical motion
(convection). However, even in the highly unstable case with $\beta=1$,
vertical speeds are sub-dominant at $\sim 30\%$ of the horizontal
speeds. This suggests that it is not essential to consider 3D models
when considering gravitational instabilities.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.cm 0.23cm
    0.0cm]{figures/growth_visc3d}
  \caption{Growth rates from the viscous 3D eigenvalue problem (black solid
    line). The other curves are obtained from the 2D dispersion
    relation (Eq. \ref{thindisk}) but with a modified gravitational
    constant as described in Appendix \ref{3dcorr}. \label{3d_visc}}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.cm 0.23cm
    0.0cm]{figures/eigenvec_vz}
  \caption{Magnitude of vertical velocities, normalized by the
    magnitude of the total horizontal velocity, of the viscous GI in 
    Fig. \ref{3d_visc}, for three cooling times: $\beta=1,\,10$, and $100$. \label{3d_visc_vz}} 
\end{figure}