### 2016-01-28 paper/paper3d.tex
revision 3 by MKL

%\documentclass[12pt]{aastex}
\documentclass[iop, numberedappendix]{emulateapj}
%\documentclass{article}
%\usepackage{draftwatermark}
%\usepackage{paralist}
\shorttitle{Non-ideal Gravitational instability}
\shortauthors{M.-K.\ Lin \& K. M. Kratter}

\usepackage{amsmath}
\usepackage{paralist}
%\usepackage{url}
\usepackage{bm}
%\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\newcommand{\p}{\partial}
\newcommand{\zmax}{z_\mathrm{max}}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\sbar}{\bar{\sigma}} 
\newcommand{\He}{\operatorname{He}}
\newcommand{\imgi}{\mathrm{i}} 
\newcommand{\ii}{\mathrm{i}}
\newcommand{\C}{\mathcal{C}^\lambda}
\newcommand{\wop}[1]{\mathcal{V}^{(#1)}}
\newcommand{\avg}[1]{\langle{#1}\rangle}
%\newcommand{\max}[1]{\mathrm{max}({#1})}
%\newcommand{\min}[1]{\mathrm{min}({#1})}
\newcommand{\zhat}{\hat{z}}
\newcommand{\khat}{\hat{k}}
\newcommand{\rhat}{\hat{r}}
\newcommand{\qthree}{Q_\mathrm{3D}}
\newcommand{\qtwo}{Q}
\newcommand{\he}{\mathcal{H}}
\newcommand{\amp}{\mathcal{A}}
\newcommand{\dd}{\delta}
\newcommand{\real}{\operatorname{Re}}
\newcommand{\imag}{\operatorname{Im}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\wbar}{\tilde{W}}
\newcommand{\qbar}{\tilde{Q}}
\newcommand{\dbigH}{\frac{H^\prime}{H}}
\newcommand{\Rp}{\frac{\rho_0^\prime}{\rho_0}}
\newcommand{\Fp}{\frac{g^\prime}{g}}
\newcommand{\Rpp}{\frac{\rho_0^{\prime\prime}}{\rho_0}}
\newcommand{\Fpp}{\frac{g^{\prime\prime}}{g}}
\newcommand{\ils}{L_s^{-1}}
\newcommand{\ilp}{L_p^{-1}}
\newcommand{\ihp}{H_p^{-1}}
\newcommand{\ihs}{H_s^{-1}}
\newcommand{\qcrit}{Q_\mathrm{crit}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\csmid}{c_{s0}} 
\newcommand{\hiso}{h_\mathrm{iso}}
\newcommand{\Hiso}{H_\mathrm{iso}}
\newcommand{\zeus}{\texttt{ZEUS-MP }}
\newcommand{\rsph}{r_\mathrm{sph}}
\newcommand{\dvx}{\dd v_x}
\newcommand{\dvy}{\dd v_y}
\newcommand{\dvz}{\dd v_z}
\newcommand{\dbx}{\dd B_x}
\newcommand{\dby}{\dd B_y}
\newcommand{\dbz}{\dd B_z}
\newcommand{\tcool}{t_\mathrm{c}}
\newcommand{\tirr}{T_\mathrm{irr}}
\newcommand{\w}{ \widetilde{W}}
\newcommand{\dphi}{\dd \Phi}
\newcommand{\hsg}{H_\mathrm{sg}}
\newcommand{\yr}{\mathrm{yr}}
\newcommand{\qvert}{Q_\mathrm{3D}}
\defcitealias{lin12}{L12} 
\defcitealias{nelson13}{N13}
\defcitealias{barker15}{BL15}
\defcitealias{mcnally14}{MP14}
\defcitealias{goldreich67}{GS67}
\defcitealias{rafikov15}{R15}

\def \OmK {\Omega_{\rm K}}

\begin{document}

%\title{untitled linear problem}
\title{On the Gravitational stability of gravito-turbulent accretion disks}
\author{Min-Kai Lin\altaffilmark{1} \& Kaitlin M. Kratter}
\affil{Department of Astronomy and Steward Observatory, \\
University of
  Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA}
\altaffiltext{1}{Steward Theory Fellow}
\email{minkailin@email.arizona.edu}

\begin{abstract}
  Low mass, self-gravitating accretion disks admit quasi-steady,
  `gravito-turbulent' states in which cooling balances turbulent viscous 
  heating. However, numerical simulations show that gravito-turbulence
  cannot be sustained beyond dynamical timescales when the cooling rate is too fast,
  or when the corresponding turbulent viscosity is too large. The result is is disk fragmentation.         
  We motivate and quantify an interpretation of disk 
  fragmentation as the inability to maintain gravito-turbulence due to  
  formal secondary instabilities driven by:  
\begin{inparaenum}[1)] 
\item 
  cooling, which reduces pressure support; and/or
\item 
  viscosity, which reduces rotational support. 
\end{inparaenum}
We analyze the gravitational stability of accretion disks 
including these non-ideal effects explicitly.    
We derive a general dispersion relation for linear waves in a 
self-gravitating, viscous, non-adiabatic disk with internal heating,
external irradiation, and cooling. We  
consider parameterized cooling functions in two and three dimensions,
as well as 
radiative diffusion in 3D. We show that 
generally there is no critical cooling rate/viscosity below which 
the disk is formally stable, although interesting limits appear 
for unstable modes with lengthscales on the order of the disk scale height.   
%when we examine the growth 
%of modes of order the disk scale height. 
We apply this new linear theory to 
global models of protoplanetary disks (PPDs) subject to
gravito-turbulence modeled as an effective 
viscosity, and cooling via dust opacity. 
We find that viscosity renders the PPD beyond $\sim 60$AU dynamically
unstable on radial %this is for Mdot = 1d-6  
lengthscales a few times the local disk thickness. This is coincident 
with the empirical condition for disk fragmentation based on a
maximum sustainable stress. 
We suggest turbulent stresses can play an active role in realistic disk 
fragmentation by removing rotational stabilization against 
self-gravity, and that the observed transition in behavior from 
gravito-turbulent to fragmenting may derive from  
instability of the gravito-turbulent state itself.
\end{abstract}

\section{Introduction}\label{intro} 
Understanding the gravitational stability of rotating disks    
is central to many astrophysical problems. In the context of 
gaseous protoplanetary disks (PPDs), gravitational instability (GI)
has two applications. It can provide gravitational torques to
transport angular momentum outwards and thus enable mass accretion
\citep{armitage10,turner14}. GI may also lead to disk fragmentation,
which has been invoked to explain giant planet formation at large
radii \citep{boss97,helled14} %% KMK need an original citation, not just review.
Studying these non-linear phenomena requires
direct numerical simulations. Nevertheless, physical insight can be    
obtained through analytical modeling.

The standard metric for the (inverse) strength of disk self-gravity is
the \emph{Toomre parameter}, 
\begin{align}\label{toomreQ_criterion}
  Q \equiv \frac{c_s\kappa}{\pi G \Sigma}  
\end{align}
\citep{toomre64}. Here, $c_s$ is the sound-speed,
$\kappa$ is the epicyclic frequency (which equals the rotation
frequency $\Omega$ in a Keplerian disk), $\Sigma$ is the surface
density and $G$ is the gravitational constant. The Toomre parameter is a
measure of the destabilizing effect of self-gravity ($G\Sigma$) against 
the stabilzing effect of rotation ($\kappa$) and pressure 
($c_s$). This is evident from the dispersion relation, 
\begin{align}\label{classic_dr}
  s^2 =  2\pi G\Sigma|k| - \kappa^2 - c_s^2k^2,   
\end{align}
which relates the growth rate $s$ and radial wavenumber $k$ for local,
axisymmetric waves in a two-dimensional (2D, razor-thin), inviscid and
adiabatic (including isothermal) disk. 
%Such disturbances are formally
%unstable when $Q<1$.  
When $Q<1$, there is a range of $k$ for which such disturbances are
unstable.  
Non-axisymmetric modes can develop for larger, but still order-unity
values of $Q$ 
\citep{lau78,papaloizou89,papaloizou91}. 
%strictly speaking non axi *local* modes are stable -> only transient
%growth
%true non-axi instability needs global disk  model 
%non-axi GI at slightly larger Q, but still characterized by Q

We emphasize that the oft-used dispersion relation and corresponding
Toomre parameter (Eqs. \ref{toomreQ_criterion} and \ref{classic_dr})
are derived from \emph{idealized} conditions:  
the base disk is laminar and does not experience any net thermal 
losses\footnote{Isothermal disks implicitly assume  
  heating and thermal losses are exactly balanced at all 
  times, so these do not cool in the current context.} 
(hereafter `cooling').  
However, real disks can cool, for example, %radiative, convective cooling
due to radiative losses. Accretion disks may also be turbulent, the 
dynamic and thermodynamic effect of which is often modeled through an 
effective viscosity \citep{shakura73,lin87,armitage01,rafikov15}. 
%{\bf  more refs here?} 
%in terms of AM transport and heating 
%assess GI in general

However, the Toomre parameter  
is still widely applied to viscous (turbulent), cooling
accretion disk models \citep[e.g.][]{gammie01,cossins09, kimura12}, 
despite the mismatch in the included underlying physics. 
We show that including non-ideal physics, such as cooling and
viscosity, in fact modifies the classic dispersion 
relation, and hence the condition for GI. Thus the Toomre condition
alone is insufficient to assess GI in realistic disks.

The goal of this work is to generalize the analytic treatment of disk
GI, by including cooling and viscosity, to allow 
a self-consistent discussion of GI in realistic accretion disk
models.  
%Specifically, we will examine how  cooling and viscosity can
%lead to GI even when $Q>1$, by removing pressure and rotational
%support, respectively.  
We first review in \S\ref{cool_gi} and \S\ref{visc_gi}, respectively,  
how cooling or viscosity can lead to GI even when $Q>1$, by removing
pressure or rotational support. We then 
discuss in \S\ref{frag_intro} how these effects may relate   
to the transition from self-regulated GI to fragmentation seen in
numerical simulations. The rest of this paper is laid out in
\S\ref{plan}.

%{\bf add sentence here more like s 1.4 laying
%out what you are about to talk about. at the very least a transition sentence or two?}

\subsection{Cooling-driven gravitational instability}\label{cool_gi}
Cooling reduces pressure support against self-gravity. The
cooling time  
$\tcool$ is the timescale over which the disk temperature $T$ is relaxed
to some floor value, which may be zero. 
%That is, $\p_t T =
%-(T-T_0)/\tcool$. 
 It is often written as 
\begin{align} \label{beta_def}
  \tcool = \beta \Omega^{-1},
\end{align}
where $\beta$ is the corresponding dimensionless cooling time. This
type of parameterized cooling, first applied by \cite{gammie01},
allows a range of thermodynamic responses to be explored.

In reality, PPD cooling is controlled by radiation   
from dust grains \citep{bell94,dalessio97,chiang97}. Although most
PPDs subject to GI are optically thick, cooling parameterized by the
$\beta$ model formally only captures optically thin cooling when
used in numerical simulations. It is, however, possible to modify the
standard cooling function to mimic optically-thick cooling (see,
e.g. \S\ref{2dppd}).

%%KMK 
Previous work has quantified the role of cooling primarily as a means to reduce the sound speed term
in $Q$ \citep[though see e.g.,][]{clarke07}. Here we will quantify how cooling enables GI
even when $Q>1$. For example, if perturbations can
cool to zero temperature, we find the above dispersion relation is
modified to read 
\begin{align}\label{cool_dr}
s^2 = 2\pi G\Sigma |k| - \kappa^2 - \gamma \left(\frac{\tcool s}{1 +
  \tcool s}\right)c_s^2k^2,
\end{align}
where $\gamma$ is the adiabatic index. Cooling increases the growth
rate by reducing the magnitude of the 
pressure term. In fact, for $t_c\in[0,\infty)$ a formal condition
for instability is  
\begin{align*}
  2\pi G \Sigma |k| > \kappa^2,
\end{align*}
which is what would be obtained from Eq. \ref{classic_dr} with 
pressure neglected ($c_s\to 0$). This condition does not actually
depend on the cooling time, and instability is
possible for any finite $Q$. %at sufficiently small scales 
Thus, the mere presence of cooling changes the qualitative nature of
GI compared to the simple Toomre condition. 
%singular limit

%{\bf numerical implications? if using standard beta cooling, then as
%  long as it's there then sufficiently small scales will be
 % unstbale. can access those scales with high resolution. defer to
 % discussion section}

%current understanding of frag is all due to cooling (viscosity is
%just a reflection of cooling)

\subsection{Viscosity-driven gravitational instability}\label{visc_gi}
%there is also viscous damping which stabilizes
Viscous disks can also develop GI even when $Q>1$. This is because, as 
demonstrated below, viscosity removes rotational support against  
self-gravity for long-wavelength disturbances 
\citep{lynden-bell74,willerding92,gammie96}. A similar effect   
occurs in dusty fluids where the required frictional forces are 
provided by dust-gas drag \citep{ward00, takahashi14}. 
In fact, this
is a mechanism to enhance particle clumping for planetesimational
formation %frag of dust 
\citep{youdin11}. Therefore it is not unreasonable to expect analogous
fragmentation in gaseous disks due to viscosity.

It is conventional to write the kinematic viscosity $\nu$ as   
\begin{align}\label{alpha_param}
  \nu  = \alpha \frac{c_s^2}{\Omega}, %dimensional argument 
\end{align}
where $\alpha$ is the dimensionless viscosity
coefficient \cite{shakura73}. This
parameterization can be modified to include more complex dependencies on the
fluid variables, but Eq. \ref{alpha_param} is the general form. 
%KMK I think this is sufficient here {\bf maybe no need to intro alpha param here?}

For an isothermal, viscous, self-gravitating disk in 2D,
\cite{gammie96} finds the approximate dispersion relation   
\begin{align}
s \simeq \frac{\nu k^2\left(2\pi G\Sigma |k| -
  c_s^2k^2\right)}{\kappa^2 + c_s^2k^2 - 2\pi G\Sigma|k|}. 
\end{align}
Assuming $Q>1$, instability occurs if 
\begin{align*}
  2\pi G \Sigma > c_s^2|k|,
\end{align*}
provided that $\nu k^2\neq0$. This condition for viscous GI is 
identical to what would be obtained from Eq. \ref{classic_dr} with
rotation neglected ($\Omega\to 0$). That is, a classically stable
disk can be destabilized by viscosity as it effectively eliminates
rotational stabilization \citep{lynden-bell74}.

%turbulence <-> viscosity assumption 
For gaseous accretion disks, a Navier-Stokes viscosity is often
implemented (as above) to mimic hydrodynamic or magneto-hydrodynamic
(MHD) turbulence \citep{shakura73}. How well turbulence can be modeled
as an effective viscosity is a separate issue 
\citep{balbus99}. However, it is reasonable to assume that these
mechanisms, which are observed in simulations  
to transport angular momentum outwards (and mass inwards),
frustate rotational support.

In this work, we use viscosity to model two possible
physical effects of turbulence: heating via dissipation and angular momentum transport. 
%KMK
We emphasize that our model for GI enabled by viscosity (hearafter
viscous GI) does not assume a particular origin for the viscosity. GI
itself or other forms of magnetic or hydrodynamic turbulence are 
allowed in this framework. We will generalize the theory of viscous GI
to include an energy equation  with viscous heating, irradiation,
explicit cooling, as well as three-dimensionality (3D) in order to
consider viscous GI in more realistic PPD models.

%effect of viscosity 
%secular instability 
%but adding viscosity means viscous heating should also be included

\subsection{Relevance to gravito-turbulent disk fragmentation}\label{frag_intro}
We develop a general framework for viscous disks without  
assuming a specific origin of the turbulence that this viscosity 
represents. 
We will, however, apply the theory to `gravito-turbulent'
disks in which the viscosity is associated with some underlying
(classic) GI, as described below.

Consider an initially laminar, $Q\gg1$ disk, without an external
heating mechanism, as it cools. The disk temperature will decline
until $Q=O(1)$, whence non-axisymmetric modes grow and heat the disk
through the dissipation of spiral shocks \citep{cossins09}. 
%new rafikov 16  
This setup permits a quasi-steady, turbulent state with
$Q =O(1)$ in which cooling is balanced by shock heating to maintain
thermal equilibrium \citep{gammie01}. 
If one represents this turbulent heating by a viscous heating, then
the $\alpha$ and $\beta$ parameters defined above will  
be inversely related (see, e.g. Eq. \ref{alpha_beta_relation}).

Numerical experiments, however, show that if the cooling time is too
small (or the viscosity is too large), say,   
\begin{align}\label{frag_cond}
  \beta < \beta_\mathrm{c} \quad
  \left(\alpha>\alpha_\mathrm{c}\right), 
\end{align}
then the disk fragments \citep{gammie01,rice05,rice11}. In fact, in 
the absence of global effects, numerical simulations show that there are
two --- and only two --- possible outcomes for self-gravitating,
cooling disks: gravito-turbulence or fragmentation. 
%more accurately: given a total simulation time, there are two
%outcomes 
%sufficiently long sims -> always frag?

There is considerable debate on the exact value of $\beta_\mathrm{c}$  
and whether or not a critical cooling time can be defined at all
\citep{meru11,lodato11,meru12,paardekooper12,hopkins13}. There are, in
addition, numerical convergence issues when simulating disk  
fragmentation, which we discuss in \S\ref{summary}.

However, it is generally accepted that \emph{steady, gravito-turbulent
  disks do not  
  exist for sufficiently rapid cooling or large viscosity}.  
This is intriguing because, as highlighted in
\S\ref{cool_gi}---\ref{visc_gi}, cooling or viscosity can reduce 
gravitational stability independently of their influence on the exact
value of the classic $Q$. This motivates a physical      
interpretation of disk fragmentation as the inability to maintain a
gravito-turbulent state due to 
secondary instabilities driven by cooling and/or viscosity.

Indeed, we find that the region where a gravito-turbulent PPD is
dynamically unstable agrees with the current consensus on  where, and
the conditions under which, a PPD is expected to fragment based on
Eq. \ref{frag_cond}, with typical critical values cited in the
literature \citep[e.g. $\beta_c\sim
  3,\,\alpha_c\sim0.1$,][]{gammie01,rice05}.

\subsection{Plan}\label{plan}
%plan of this paper 
%aim to analyse GI in realistic disks with minimal input from numerical
%simulation

The basic equations, disk equilibria, cooling and viscosity models are 
given in \S\ref{basic}. The linear stability problem is defined in 
\S\ref{linear}. We present results with parameterized `beta' cooling in 
\S\ref{result_2d} and \S\ref{3ddisk} for 2D and 3D disks,
respectively. %key result: generalized dispersion relation 
In \S\ref{2dppd} we consider PPDs with realistic viscosity/cooling   
models, including radiative diffusion in 3D, and determine where PPDs
are gravitationally unstable and why. We summarize our results in
\S\ref{summary} with a discussion of how our models may aid the 
physical understanding of fragmentation in realistic PPDs.

\input{model3d}
\input{linear3d}
\input{results2d}
\input{results3d}
\input{application} 
\input{summary}

%\acknowlegement
%S. Stahler 
%A. Youdin

\appendix
\input{appendix3d}

\bibliographystyle{apj}
\bibliography{ref}

\end{document}