### 2016-02-02 paper/application.tex
submitted version

\section{Application to protoplanetary disks}\label{2dppd}
We now apply our linear framework to assess the stability of  
PPDs. We consider the gravito-turbulent disk models recently
developed by \citet[][hereafter \citetalias{rafikov15}]{rafikov15}. 
This 2D, Keplerian disk orbits a Solar 
mass star and is defined by the following parameters,

\begin{itemize}
  \item $\dot{M}$, the global radial mass accretion rate;
  \item $Q_0$, the value of the 2D Toomre parameter where the disk is
    gravito-turbulent;
  \item $T_\mathrm{irr}$, the irradiation temperature;
  \item $\alpha_m$, the dimensionless viscosity associated with other
    sources of turbulence, such as magneto-rotational instabilities
    (MRI, see also \S\ref{visc_gi} and \S\ref{MHD}). 
\end{itemize} 
These properties serve as inputs for calculating thermal equilibrium, and mass and
angular momentum conservation. Together with an opacity law, 
they allow us to construct a global disk model with surface density $\Sigma(R)$ and
temperature $T(R)$ where $R$ is the global cylindrical radius from the
star. \citepalias[See][for details.]{rafikov15} These profiles give $Q(R)$,
required for input into our linear framework. We derive other
dimensionless parameters below.

Although our disk is global in extent, here we consider the local
stability at each radius. We are thus neglecting any global
instabilities \citep{adams89,lodato05,kratter10} as well as any
evolution of the disk properties in response to GI.

% We are thus neglecting the slow, background viscous radial flow; as well
%as any global gravitational instabilities \citep{lodato05}. {\bf more refs?}

\subsection{Effective $\alpha$}
In a steady, viscously accreting Keplerian disk 
with constant $\dot{M}$ we have,
approximately, $\nu\Sigma = \dot{M}/3\pi$. Hence
\begin{align}
  \alpha = \frac{\dot{M}}{3\pi}\frac{\Omega(R)}{c_{s}^2(T)\Sigma(R)},  
\end{align} 
which sets the viscosity coefficient to be used at each radius. Note that the disk profiles 
employed here give constant $\dot{M}$ rather than constant $\alpha$.

\subsection{PPD beta cooling}\label{ppd_cooling}
Energy loss in \citetalias{rafikov15} is given by 
\begin{align}\label{real_cool}
  \Lambda = \frac{2\sigma}{f(\tau)}\left(T^4 - T_\mathrm{irr}^4\right)
\end{align}
per unit area, where
\begin{align}
  f(\tau) = \tau + \frac{1}{\tau}, \label{ftau} 
\end{align}
and
\begin{align}
  \tau = \kappa_d(T)\Sigma
\end{align}
is the optical depth. Recall Eq. \ref{opacity_law} 
is our opacity model where $\kappa_d\propto T^b$ with $b=2$. 
Eq. \ref{ftau} accounts for cooling in the optically-thin ($\tau\ll
1$) and optically-thick ($\tau\gg1$) regimes.

Note that Eq. \ref{real_cool} falls within our definition of a beta
cooling prescription, because 
it is an explicit function of the thermodynamic states. However, we
formulated the linear problem with the standard beta cooling function 
given by Eq. \ref{beta_cool}, which has a different (less realistic)
dependence on disk temperature. 
 In order to adapt the existing  framework to the above PPD cooling function, we 
need to identify the equivalent $\beta$ and $\theta$ parameters that are required for the 
linearized equations (see \S\ref{linear_bcool}).

%Since we are interested in the linear response of the system, the
%corresponding cooling parameters $\beta$ and $\theta$ are found by
%comparing the linearized form of the cooling function
%(Eq. \ref{real_cool}) to that of the model function
%(Eq. \ref{beta_cool}).

%{\bf as opposed to comparing their unperturbed forms, as done in literature}
Linearizing Eq. \ref{real_cool} gives   
\begin{align}\label{linear_real_cool}
  (\gamma-1)\frac{\delta\Lambda}{\Sigma} = \frac{2\sigma(\gamma-1)
    T^4C_1}{f(\tau)c_{s}^2\Sigma}\left(\frac{\delta P}{\Sigma} -
  \frac{C_2}{C_1}c_{s}^2\frac{\delta\Sigma}{\Sigma}\right), 
\end{align}
where
\begin{align}
C_1(\tau, T) &= 4 - b\times g(\tau, T),\\ 
C_2(\tau, T) &= 4 + (1-b)\times g(\tau, T),\\ 
  g(\tau, T) &= \left( \frac{\tau^2-1}{\tau^2+1}\right)\left(1 -
  \frac{T_\mathrm{irr}^4}{T^4}\right). \label{g_def}
\end{align}
Comparing Eq. \ref{linear_real_cool} with the linearized form of the
standard cooling function, Eq. \ref{linear_beta}, we identify
\begin{align}
  &\beta = \frac{f(\tau)c_s^2\Sigma\Omega}{2\sigma(\gamma-1)C_1T^4},\label{real_beta}\\
  &\theta = \frac{C_2}{C_1},\label{real_theta} 
\end{align}
to be used in the 2D dispersion relation (Eq. \ref{thindisk}). 
Eq. \ref{real_beta} represents a physical cooling time \emph{for the
  perturbations}, and is consistent with previous definitions within
factors of order unity \citep[e.g.][their Eq. 2]{kratter10}.

%KMK
%However, $\theta$ is no longer a direct proxy for the irradiation level
%as it was in the standard beta cooling model.
%is now an effective irradiation level. 
Eq. \ref{real_theta} shows that $\theta$ 
is related to the true irradiation temperature $\tirr$ through the function $g$ given by 
Eq. \ref{g_def}, and is therefore a only a weak function of the irradiation temperature. 
%A physically non-irradiated disk with $\tirr=0$ will 
%have a non-zero value of the `irradiation parameter', $\theta\neq0$. 
More specifically $\theta = O(1)$ 
for all $\tirr <T$, and for our adopted opacity law, Eq. \ref{opacity_law}, 
\begin{align*}
  \theta = \frac{4-g}{4-2g}. 
\end{align*}
Thus for $T_\mathrm{irr} = 0$ we have $5/6<\theta<3/2$ by considering
$\tau\to 0,\,\infty$. However, for $T=\tirr$ we have $\theta = 1$, as
expected intuitively.    
%%KMK 
%This results from the fact that PPD cooling generally depends on both
%$P$ and $\Sigma$, but in the absence of irradiation standard beta
%cooling only depends on $P$. 
%{\bf KMK: we should discuss}

%On the
%other hand, fully irradiated disks with  $T=T_\mathrm{irr}$ have 
%$\theta=1$, as is in the previous problem formulation
%(Eq. \ref{tirr_def}).

\subsection{Inviscid stability condition}
%We note that for our adopted opacity law, Eq. \ref{opacity_law}, 
%\begin{align*}
%  \theta = \frac{4-g}{4-2g},
%\end{align*}
%implying $5/6<\theta<3/2$. 
With our new linearized cooling function in hand, from the discussion
in \S\ref{2d_inviscid} and by applying 
Eq. \ref{stable_condition},  we conclude that without viscous effects 
the disk is stable everywhere if  
\begin{align} 
  \gamma > \frac{3}{2} \quad \text{and} \quad Q >
  \sqrt{\frac{6}{5}} \label{ppd_invisc_cond} 
\end{align} 
are both satisfied.

PPDs become irradiation-dominated at large distances from the star,
where $T\to\tirr$ and $\theta\to 1$ \citep{dalessio97,kratter11}.
Then Eq. \ref{ppd_invisc_cond} relaxes to 
$\gamma,\, Q > 1$ in the outer disk. The condition on $\gamma$
is then guaranteed. On the other hand, numerical 
simulations of gravito-turbulence show that $1\lesssim Q \lesssim 2$
\citep{gammie01,rice11}, and the second inequality is generally
satisfied.   
%KMK
Taken together, this suggests that in the outer regions of a realistic
PPD, cooling %by itself may not be not responsible 
may not be the primary cause for a secondary 
instability of a gravito-turbulent disk, leading to
fragmentation. This leaves viscous GI as the only possible culprit
within our framework, as we illustrate below.

\subsection{Example 2D calculation}\label{pp2d_example}
We relax the inviscid assumption and consider a fiducial disk  model with
$\dot{M} = 10^{-6}M_\sun\,\yr^{-1}$, $Q_0=1.5$,  
$\tirr=10\mathrm{K}$, and $\alpha_m=10^{-3}$. 
%KMK
Such a high accretion rate is consistent with those expected for 
young protostellar disks \citep{shu77,enoch09}. Similarly,
$10\mathrm{K}$ is a conservatively low background irradiation level
consistent with cloud temperatures in star forming regions
\citep{plume97,johnstone01}. 
Stellar irradiation will typically elevate $\tirr$ in addition to
adding a radial dependence \citep{kratter08} 
We adopt the opacity scale $\kappa_{d0} =
5\times10^{-4}\mathrm{cm}^2\,\mathrm{g}^{-1}\,\mathrm{K}^{-2}$  as in
\citetalias{rafikov15}. We use $\gamma=1.6$, approximately applicable
to an ideal molecular gas at low temperatures 
%KMK this is just true...and used by other authors
%\citep{rice11,baehr15}. 
This choice of $\gamma$ satisfies the global inviscid stability condition
(Eq. \ref{ppd_invisc_cond}). %which simplifies the problem
%for outer disk choice of gamma doesn't matter 
%avoid complications from over-stable thermal modes 
Fig. \ref{rafikov_model} shows the equilibrium 
disk profile in terms of $Q$, $\alpha$, $\beta$, and $\theta$.  
These profiles serve as input to the 2D dispersion relation
(Eq. \ref{thindisk}, Eq. \ref{bigA}---\ref{bigF}).

Fig. \ref{rafikov_growth} shows growth timescales and
optimum wavenumbers for viscous GI in this fiducial model. For
comparison we also plot a case with lower accretion rate, 
$\dot{M}=10^{-7}M_\sun\,\yr^{-1}$; and analytic estimates based on
Eq. \ref{gammie_smallk} (instead of Eq. \ref{gammie_maxrate_simple}
since here $\theta\sim 1$) which gives the optimum wavenumber and
growth rates as 
\begin{align}
  |K| = \frac{3}{2\theta Q}, \quad
%\end{align}
%with growth rate
%\begin{align} 
  S = \frac{27\alpha}{16\theta^3Q^4}. 
\end{align}
These are similar to the isothermal results of
\citet[][their Eq. 19 and 21, respectively]{sterzik95}, and identical if one
takes $\theta=1$. %showing that irradiation makes the disk behave like
                  %isothermal disk

The most unstable wavelength is a few times the disk thickness. 
%KMK
So long as $H\ll R$, this result is consistent with our use of the
local approximation. For our fiducial disk, $H/R \sim 0.07$ around 
$R\sim 100$AU, and $H/R <0.25$ throughout the disk. 
The increase in $|K|$ from 
$\sim 10\mathrm{AU}$ to $\sim 20\mathrm{AU}$ is due to the decrease in
$Q$, while that from $\sim 
60\mathrm{AU}$ to $\sim 100\mathrm{AU}$ occurs as the disk transitions
from the 
optically-thick to optically-thin regime. The mismatch between the
numerical and analytic solutions at large
distances is expected since the above expressions assume $|K|\ll
1$. Nevertheless the analytic estimates reproduce qualitatively correct
behavior.

The fiducial disk is subject to viscous GI on dynamical timescales
($\lesssim 10$ orbits) for $R\gtrsim60$AU. We note this transition
radius is also implied by Eq. 16 in \cite{kratter10}. Coincidentally,
beyond this radius $\alpha\gtrsim 0.1$ (and 
$\tcool\Omega\lesssim 3$), which is often quoted as a condition for disk
fragmentation   
\citepalias[e.g.][]{rafikov15}. Thus viscous GI may be responsible for 
the transition between gravito-turbulence and fragmentation due to the
removal of rotational support by viscous (turbulent) stresses.

On the other hand, the lower $\dot{M}$ model also attain $\tcool\lesssim
3\Omega^{-1}$ beyond $\sim 60$AU, but $\alpha\lesssim 0.03$
everywhere. 
Applying empirical cooling conditions for fragmentation may then lead to
contradiction. Instead, if viscosity is the physical cause for
fragmentation, then our result suggest the lower $\dot{M}$ disk should
not fragment (at least much less likely than our fiducial case) 
because the instability cannot develop on orbital timescales.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
    0.0cm]{figures/ppd_2d_basic}
  \caption{Equilibrium profile obtained from the disk model developed
    by \cite{rafikov15}, with parameters $\dot{M} =
    10^{-6}M_\sun\yr^{-1}$, $Q_0=1.5$, $\tirr=10\mathrm{K}$, and
    $\alpha_m=10^{-3}$.   
    \label{rafikov_model}}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 2cm 0cm
    0.0cm]{figures/ppd_2d_growth}\\
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
    0.cm]{figures/ppd_2d_maxk}
  \caption{Black lines show growth timescales (top) of the most
    unstable wavenumber (bottom) for viscous, 
    self-gravitational modes in the 2D PPD model shown in
    Fig. \ref{rafikov_model}. Blue curves are for the same disk
    model but with a lower accretion rate. Solid curves are obtained
    numerically from Eq. \ref{thindisk}, and dotted curves 
    are analytic results based on Eq. \ref{gammie_smallk}. 
    \label{rafikov_growth}}
\end{figure}

\subsection{3D PPD with radiative diffusion}
%appropriate for not-so-large distances where irrad can be neglected 
We briefly consider 3D PPDs with explicit radiative diffusion
(\S\ref{rad_cool}). 
Given the $\alpha(R)$ and $Q(R)$ profiles obtained
from the 2D model above, at each radius $R$ we obtain the vertical
structure from Eq. \ref{vert_eq1}---\ref{thermal_eq}, with 
Eq. \ref{rad_cool1}---\ref{rad_cool2} for the radiative flux.
We then solve the 3D eigenvalue problem as in \S\ref{3ddisk}, with the
additional boundary condition that the disk surface temperature is
fixed, $\delta T(\zmax) = 0$. 
We use the fiducial disk model as in \S\ref{pp2d_example} but 
with $T_\mathrm{irr}=0$, since our simple radiative diffusion
treatment does not include irradiation (\S\ref{rad_cool}). %This also
                                %limits us 
%to smaller radii than before.  
%physically, irrad is negligible for opt thick regions, and we are
%restricted to such regions if applying rad diff
We use a slightly smaller vertical
domain with $\rho(\zmax)=0.1\rho_0$.  %to avoid very low density

Fig. \ref{rafikov_growth3d} shows the growth rates and most unstable 
wavenumber for $R\in[10,100]$AU; along with the corresponding 2D 
results matched with softened self-gravity. There is good agreement for
$R\lesssim60$AU where the disk is optically thick ($\tau\gtrsim
1$) and thus both models apply. However, beyond $60$AU where the disk
becomes optically-thin, radiative diffusion (the 3D curve) is not
valid and under-estimates the growth rates. %; while the $f(\tau)$
%function approximately captures opticall-thin cooling in the 2D case.  
Nevertheless, the transition radius of  
$\sim60$AU, beyond which growth timescales become dynamical,    
can be correctly calculated within the 2D framework.  
%notice the small softenings used 
%This is not surprising since the
%instability is fundamentally driven by viscous and self-gravity forces in the
%momentum equations, not the energy equation.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 2cm 0cm
    0.0cm]{figures/ppd_3d_rates}\\
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
    0.8cm]{figures/ppd_3d_maxk}
  \caption{Growth timescales (top) of the most unstable
    wavenumber (bottom) for viscous GI in a 3D PPD with radiative
    diffusion  (black lines). Asterisks are corresponding results
    obtained from the corresponding 2D problem with softened gravity.   
    % The black curves are obtained
  %  from the 3D eigenvalue problem with radiative diffusion,
   % applicable to optically-thick regions ($R\lesssim 60$AU); 
   % while the green curves are obtained from the 2D corresponding
   % model, which is applicable to both optically-thick and
   % optically-thin regimes.  
    \label{rafikov_growth3d}}
\end{figure}